### 2024-03-29 paper/paper.qmd
Adding simulate data, test data, clean data, models, and the rough copy paper.

---
title: "NFL Analysis"
author: "Alexander Guarasci"
thanks: "Code and data underpinning this paper are available at: https://github.com/AlexanderG123/nfl_analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`" # This will add the current date
date-format: long
bibliography: references.bib
execute:
  echo: false
format: pdf
fig-pos: "H"
abstract: |
 This paper is the shit

---

```{r}
knitr::opts_chunk$set(echo = FALSE, results = "hide", message = FALSE, warning = FALSE)

```

# Introduction {#sec-intro}

When it comes to professional football, there has long been conjecture that a team's performance in games is closely correlated with their offensive yardage total, which includes both passing and running yards. In order to find patterns and insights that could affect team strategy and fan comprehension of the game dynamics, this study conducts a thorough statistical analysis of the complex link between various offensive yard measurements and the probability of winning NFL games. This research attempts to quantify the effect of offensive performance on game outcomes by methodically analyzing play-by-play data for the 2023 NFL season. It provides a thorough summary of how offensive yardage affects a team's victory both throughout the entire game and when focusing only on the statistics for the fourth quarter.

The constant discussion among coaches, commentators, and fans over the best tactics for winning football games is what inspired this research. Although conventional wisdom has frequently emphasized the significance of a balanced offence—one that features a dynamic rushing (note: rushing and running are used interchangeably in this paper) attack in addition to a solid passing game—the empirical evidence that backs up this claim has been largely anecdotal. This paper presents a data-driven investigation of the correlations between various offensive methods and winning games. This analysis will help to provide insights into tactical choices and advance our knowledge of football analytics.

This analysis was conducted through the R programming language [@citeR] and supplemented with the analytical tools Tidyverse [@citetidyverse] and caret [@citecaret]. The dataset was assembled from play-by-play records of NFL games from NFLVerse [@citenflverse]. The data is analyzed using statistical methods, such as logistic regression, which provides insights into the relative value of passing and rushing in securing game wins. This method opens the door for further study in sports analytics and offensive strategy optimization while also illuminating the prevailing dynamics of effective offensive methods.

This study aims to make a substantial contribution to the conversation about sports analytics by offering a strong analytical framework for examining the connection between NFL game-winning and yards. It reveals insights into offensive strategy through in-depth statistical analysis and model-building, providing information for analysts, players and fans. The results of this study show how important data analysis is to sports strategy and how quantitative methods can potentially help improve competitive advantages in professional football.

# 2 Data {#sec-data}

The data used in this paper was taken from nflverse [@citenflverse], which consists of “a set of packages dedicated to data of the National Football League.”.

# 2.1 The Dataset

The foundation of the dataset that was ultimately used in this paper is the play-by-play data from across the 2023 NFL season. This includes an unbelievably dense array of information from every snap over the entirety of the season consisting of 372 variables. These include: seconds left in the half, touchdown probability, expected points added, the probability of taking a safety, the position the snap was taken in, a description of the play, and many, many more. In order to filter this information the play-by-play data was cleaned and new variables were created.

Other statistics that were taken into consideration but ultimately not used included more detailed play-by-play assessments and analytics on the performance of individual players. It was determined that the comprehensive dataset with an emphasis on team-level performance indicators was the most suitable for investigating the overall correlation between offensive yardage and game results. This method avoids player-specific variability and enables a macro-level analysis, which is in accordance with the study's goal of outlining general strategic insights.

# 2.2 Variables of interest

This comprehensive dataset encapsulates a wide array of variables, central to which are “RushingYards”, “PassingYards”, “TotalYards”, “home_score”, “away_score”, “Winner”, “Win”, “Q4_RushingYards”, “Q4_PassingYards” and “Q4_TotalYards”. RushingYards and  PassingYards represent the yards gained by a team through rushing and passing plays, respectively, within a single game, while Q4_RushingYards and Q4_PassingYards represent the rushing and passing yards in the fourth-quarter. “RushingYards” is taken as the sum of the variable “rushing_yards” from the original dataset with the same “game_id” (so that we get the rushing yards from each game) and “posteam” (so each team that plays gets has their rushing yards tracked).  PassingYards, Q4_RushingYards and Q4_PassingYards are similarly constructed. TotalYards is a constructed variable, representing the sum of RushingYards and PassingYards, devised to encapsulate a team's overall offensive performance. Q4_TotalYards is constructed as the sum of Q4_RushingYards and Q4_PassingYards. The scores of the home and away teams are captured by home_score and away_score, respectively, which is data that is provided and that does not need to be constructed. The Winner is determined based on the comparison of these scores, classified as the team that won (for example, “Bal” if the Baltimore Ravens won). Finally, the binary Win variable signifies the outcome from the perspective of the team being analyzed (1 for a win, 0 for a loss, and “TIE” for a tie, but because there were no ties last season there are no instances of “TIE”).

Summary statistics were generated to provide an overarching view of the data's distribution and central tendencies. These statistics reveal an average of approximately 113 rushing yards and 238 passing yards per game, with total yards averaging around 350. The distribution of these variables, along with the game outcomes (Win), was visually examined through histograms and box plots, elucidating the central tendencies and variability within the data. These visual explorations further affirmed the preliminary assumption that a higher aggregate of offensive yards tends to correlate with a higher likelihood of winning.

The measurement aspect of this data is something that can be scrutinized. If the reader of this paper has ever watched an NFL game, they are likely very aware of the lack of certainty with regard to where the football is placed.  A player gains yards until they are tackled. Following the tackle the player hands the ball to the referee who then places it where he or she believes the ball was when the player went down, and which is where the next play begins. This spot is where the data is recorded.  Although this is a severely flawed technique that often infuriates viewers, it is not something that concerns our analysis because we are curious about the end results, how each gained yard benefits a team. We do not care about this difference because it does not impact the results of a game. For example, if a player gets six yards on a play but the referee only spots him three yards, the end result is a three-yard gain, the six yards will have no impact on the outcome.

# 3 Models {#sec-Models}

Two logistic regression models were used in our study to analyze the NFL game dynamics, with a particular focus on the relationship between a team's cumulative running and passing yards and its likelihood of winning. Both the full game and just the fourth quarter are catered to by these models, which provide a multi-layered understanding of strategic applications at different points of the game.

# 3.1 Model 1: Overall Game Analysis

The first logistic regression model scrutinizes the entire game, using rushing and passing yards as predictor variables to estimate the odds of winning. This model, expressed as $$
\log\left(\frac{p_1}{1 - p_1}\right) = \beta_0 + \beta_1 \cdot \text{RushingYards} + \beta_2 \cdot \text{PassingYards}
$$
, where p1 denotes the probability of winning, facilitates a broad analysis of strategic effectiveness. The coefficients, representing rushing and passing yards respectively, illuminate the value of each yard gained in the context of winning. The positive coefficient for rushing yards implies a direct correlation between a stronger rushing offence and the likelihood of victory, underscoring the strategic advantage of a balanced offensive game. This model is apt for providing a holistic view of game strategy, allowing analysts and teams to gauge the overall impact of their offensive efforts throughout the match.

# 3.2 Model 2: Fourth Quarter Focus

The second logistic regression model narrows the focus to the fourth quarter, a critical period where games are often decided. By analyzing the same predictors but confined to the final quarter, this model, $$
\log\left(\frac{p_2}{1-p_2}\right) = \beta_3 + \beta_4 \cdot \text{Q4\_RushingYards} + \beta_5 \cdot \text{Q4\_PassingYards}
$$
, offers insights into the tactical shifts that dominate the endgame. Unlike the overall game model, the fourth-quarter model might reveal a nuanced strategic landscape, such as a negative coefficient for passing yards, indicating the potential pitfalls of a pass-heavy approach under pressure. This focused analysis is instrumental in understanding the dynamics of clutch gameplay, shedding light on the adjustments teams make in response to the scoreboard and the clock.

# 3.3 Justification and Appropriateness

The choice of logistic regression is prudent in light of the binary structure of the outcome variable (win or loss) and the models' ability to convey the connection between playing strategies (rushing and passing) and winning odds. The interpretability of logistic regression emphasizes its suitability even more; the exponential of the coefficients accurately measures the change in odds that come with every extra yard gained, whether by passing or rushing. This gives you a concise, useful grasp of how various aspects of the game affect your chances of winning.

Furthermore, the choice to use two different models takes into account the intricate, multidimensional character of football strategy. The fourth-quarter model focuses on the crucial choices and plays that frequently determine the result, while the overall game model reflects the persistent efforts and strategies that mould the game's early and middle phases. This dual-model method enables a thorough analysis that acknowledges the urgency of the game's closing moments as well as the marathon nature of establishing a lead.

In summary, the use of logistic regression models to analyze NFL game outcomes offers a robust methodological framework for dissecting the strategic components of football. By evaluating the entire game and the pivotal fourth quarter separately, these models provide nuanced insights that can guide teams in refining their strategies for sustained success and clutch performance.

# 4 Results: {#sec-results}

As an overview of the whole data set, of the 285 games played in the NFL regular season and playoffs, there were 90 games where a team rushed for fewer yards than their opponent and won while there were 109 games where a team threw for fewer yards than their opponent and won. Prioritizing the rushing attack is further bolstered by the model which suggests that a rushing yard is more valuable than a passing yard. In only 72 games a team rushed and passed for more yards than their opponent and won the game and in 14 games a team rushed for and passed for fewer yards while winning the game.

# 4.1 Graphical Results
```{r}
#| label: fig-1
library(ggplot2)
library(tidyverse)
data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
ggplot(data, aes(x = RushingYards, y = PassingYards, color = as.factor(Win))) +
  geom_point() +
  labs(title = "Rushing Yards vs. Passing Yards by Win/Loss",
       x = "Rushing Yards",
       y = "Passing Yards",
       color = "Outcome") +
  theme_minimal()
```
@fig-1 shows the rushing yards and passing yards of winning teams over the 2023 NFL season. In examining the graph it is clear that the overall trend, more often than not, is that more rushing yards and more passing yards increase the likelihood of winning. The biggest outlier game, which resulted in the Miami Dolphins Rushing for 350 yards and passing for 376 yards on the Denver Broncos, led to a 70 to 20 victory for the Dolphins. Prior to that game, no team in the 21st century had scored 70 points in a game, this anecdotally shows how being able to run and pass the ball effectively is exceedingly helpful when it comes to scoring points and ultimately winning ball games.

```{r}
#| label: fig-2
data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
ggplot(data, aes(x = as.factor(Win), y = TotalYards, fill = as.factor(Win))) +
  geom_violin()+
  labs(title = "Distribution of Total Yards by Win/Loss",
       x = "Outcome (0 = Loss, 1 = Win)",
       y = "Total Yards",
       fill = "Outcome") +
  theme_minimal()
```
@fig-2 supports this idea, it is a violin plot of the total yards of winning and losing teams (total yards is the sum of passing yards and rushing yards). What it shows is that winning teams never had fewer than 200 total yards, and had more total yardage on average than the teams that lost. It also shows that most winning teams had around 400 total yards while losing teams had just over 350 yards.

```{r}
#| label: fig-3
data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
avg_yards_by_outcome <- data |>
  group_by(Win) |>
  summarize(AvgRushingYards = mean(RushingYards, na.rm = TRUE),
            AvgPassingYards = mean(PassingYards, na.rm = TRUE)) |>
  pivot_longer(cols = -Win, names_to = "YardType", values_to = "AverageYards")

ggplot(avg_yards_by_outcome, aes(x = as.factor(Win), y = AverageYards, fill = YardType)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("AvgRushingYards" = "skyblue", "AvgPassingYards" = "orange")) +
  labs(title = "Average Rushing and Passing Yards for Wins and Losses",
       x = "Outcome (0 = Loss, 1 = Win)",
       y = "Average Yards",
       fill = "Type of Yards") +
  theme_minimal()
```
This is further reinforced by @fig-3 which shows the average passing and rushing yards of winning and losing teams. As you can see, the winning teams rushed and passed for more yards. The results become more interesting when we break them down in the fourth quarter.

```{r}
#| label: fig-4
avg_q4_yards_by_outcome <-  read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv") %>%
  group_by(Win) %>%
  summarize(AvgRushingYards = mean(Q4_RushingYards, na.rm = TRUE),
            AvgPassingYards = mean(Q4_PassingYards, na.rm = TRUE)) %>%
  pivot_longer(cols = -Win, names_to = "YardType", values_to = "AverageYards")

# Plot the average Q4 rushing and passing yards for wins and losses
ggplot(avg_q4_yards_by_outcome, aes(x = as.factor(Win), y = AverageYards, fill = YardType)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("AvgRushingYards" = "skyblue", "AvgPassingYards" = "orange")) +
  labs(title = "Average Q4 Rushing and Passing Yards for Wins and Losses",
       x = "Outcome (0 = Loss, 1 = Win)",
       y = "Average Yards in Q4",
       fill = "Type of Yards") +
  theme_minimal()
```
@fig-4 shows a stark difference in pass yards and rushing yards of the losing team.  This trend is substantially more pronounced than @fig-3 is and provides a much more interesting backdrop, teams that lose football games pass the ball substantially more than the teams that win, and in particular they do so in the fourth quarter.

```{r}
#| label: fig-5
q4_final_data_for_modeling <-  read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv") 
#4th q data Q4_RushingYards,Q4_PassingYards
ggplot(q4_final_data_for_modeling, aes(x = Q4_RushingYards, y = Q4_PassingYards, color = as.factor(Win))) +
  geom_point() +
  labs(title = "Rushing Yards vs. Passing Yards by Win/Loss 4th Q",
       x = "Rushing Yards",
       y = "Passing Yards",
       color = "Outcome") +
  theme_minimal()

```

```{r}
#| label: fig-6
data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
ggplot(data, aes(x = Q4_RushingYards, fill = as.factor(Win))) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of Rushing Yards by Win/Loss",
       x = "Rushing Yards",
       y = "Density",
       fill = "Outcome") +
  theme_minimal()
```
```{r}
#| label: fig-7
data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
ggplot(data, aes(x = Q4_PassingYards, fill = as.factor(Win))) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of Passing Yards by Win/Loss",
       x = "Rushing Yards",
       y = "Density",
       fill = "Outcome") +
  theme_minimal()
```
Comparing @fig-5 with @fig-1, there is quite a contrast.  Whereas @fig-1 showed that winning teams ran and threw the ball more, @fig-5 shows the winning teams hugging the x-axis (more fourth-quarter rushing yards and fourth-quarter fewer passing yards), with losing teams hugging the y-axis (fewer fourth-quarter rushing yards and more fourth-quarter passing yards).  @fig-6 and @fig-7 show the distributions of fourth-quarter rushing and passing yards by winning and losing teams and both show how pronounced this trend is.

# 4.2 Model Results

This paper relied on two models. The first was a logistic regression of overall game data while the second was a logistic regression on purely fourth quarter data.

The regression on overall data indicated a positive relationship between rushing yards and passing yards and winning the game, with rushing yards being considered more valuable with regard to winning.  This model was able to correctly predict 65.79% of outcomes in the test data, predicting 71.93% of actual wins and 59.65% of actual losses.  It demonstrated a meaningful ability to predict game outcomes based on rushing and passing yards but there was room for improvement.

The second regression, using only fourth-quarter data, had more interesting results. Similar to the first model, there was a positive relationship between rushing yards and winning, but with fourth-quarter rushing yards being more valuable than overall game rushing yards.  Interestingly, there was a negative relationship between passing yardage and winning the game, in other words, the more passing yards a team had in the fourth quarter, the more likely that team was to lose the game.  This model was more accurately able to predict the outcome of games, correctly predicting 78.95% of winners and 66.67% of losers, for an overall balanced accuracy of 72.81%.

# 5 Discussion

The analysis and models presented in this paper provide a deep dive into the strategic dynamics of NFL games, with a particular focus on the impact of rushing and passing strategies on game outcomes. Through the lens of logistic regression models—both encompassing overall game performance and isolating the critical fourth quarter—we uncover nuanced insights into the traditional adage of football: "the run game wins matches." The findings affirm this notion but also highlight the complexity behind strategic decisions made on the field, especially under the pressure of the game's closing moments.

One significant insight this paper brings to light is the tangible value of a balanced attack in securing victories. The overall game model suggests a positive correlation between both rushing and passing yards with winning, reinforcing the importance of versatility in offense. This is a testament to the complexity of football strategy, where being unpredictable and capable in both facets of the game complicates the defense's task. However, the differential valuation of rushing over passing yards in the models indicates a slight strategic edge for the ground game, potentially due to the clock management and lower turnover risks associated with rushing.

The fourth-quarter model, however, opens a window into the high-stakes decision-making that defines tight contests. The negative relationship between fourth-quarter passing yards and winning underscores a common game scenario: teams behind on the scoreboard resort to the air to catch up quickly. While this strategy is rational and sometimes the only option, it is less likely to result in a win, as indicated by the model. This outcome may also reflect the defensive adjustments expecting passes, further decreasing the efficiency of a one-dimensional approach.

The study is not without its limitations. The reliance on aggregate statistics like total yards overlooks the game's situational context, such as the down-and-distance, the impact of turnovers, and special teams' play, which can significantly influence game outcomes. Moreover, the models do not account for the intricacies of clock management, a critical aspect of late-game strategy. Future research could benefit from incorporating these elements into more sophisticated models, perhaps employing play-by-play data to capture the game's flow and strategic shifts more accurately.

Looking forward, the exploration of strategic dynamics in NFL games could be enriched by integrating advanced metrics such as Expected Points Added (EPA) and Win Probability Added (WPA), which provide a more nuanced understanding of each play's impact. Additionally, qualitative analyses of coaching decisions, player performances in clutch situations, and the psychological aspects of late-game pressure could offer comprehensive insights into the art and science of winning football games. As the sport continues to evolve, so too will the strategies that define its outcomes, necessitating ongoing analysis and adaptation by teams and analysts alike.

\newpage

# References

### 2024-04-02 paper/paper.qmd
Update Read Me

@@
 ---
-title: "NFL Analysis"
+title: "Field Dynamics" 
+subtitle: "An Analysis of Offensive Strategies in the Modern Day NFL"
 author: "Alexander Guarasci"
@@
 abstract: |
- This paper is the shit
+ This study examines the strategy behind National Football League (NFL) games, focusing on how teams' offensive yardage—gained through both rushing and passing—affects their chances of winning.
+Analyzing play-by-play data from the 2023 NFL season, we discovered that while both rushing and passing yards contribute to game victories, a stronger rushing attack significantly increases a team's likelihood of winning.
+These findings highlight the critical role of offensive strategy in football, providing valuable insights for teams looking to optimize their game plans for better outcomes.
 
@@
 It provides a thorough summary of how offensive yardage affects a team's victory both throughout the entire game and when focusing only on the statistics for the fourth quarter.
+Ultimately trying to answer the questions, how do offensive yards impact the outcome of an NFL game, and what is more valuable passing yards or rushing yards[^note]?
+
+[^note]: Rushing yards and running yards are used interchangeably throughout this paper
 
 The constant discussion among coaches, commentators, and fans over the best tactics for winning football games is what inspired this research.
-Although conventional wisdom has frequently emphasized the significance of a balanced offence—one that features a dynamic rushing (note: rushing and running are used interchangeably in this paper) attack in addition to a solid passing game—the empirical evidence that backs up this claim has been largely anecdotal.
+Although conventional wisdom has frequently emphasized the significance of a balanced offence—one that features a dynamic rushing attack in addition to a solid passing game—the empirical evidence that backs up this claim has been largely anecdotal.
 This paper presents a data-driven investigation of the correlations between various offensive methods and winning games.
@@
 
-The data used in this paper was taken from nflverse [@citenflverse], which consists of “a set of packages dedicated to data of the National Football League.”.
+The data used in this paper was taken from nflverse [@citenflverse], which consists of “a set of packages dedicated to data of the National Football League.”
 
@@
 # 4.1 Graphical Results
-```{r}
+```{r, fig.cap= "The Rushing and Passing Yards of Every Game in 2023"}
 #| label: fig-1
@@
 
-```{r}
+```{r, fig.cap= "Violin Plot of Total Yards by Wins and Losses"}
 #| label: fig-2
@@
 
-```{r}
+```{r, fig.cap= "Average Yards by Wins and Losses"}
 #| label: fig-3
@@
 
-```{r}
+```{r, fig.cap= "Average Fourth-Quarter Yards by Wins and Losses"}
 #| label: fig-4
@@
 
-# Plot the average Q4 rushing and passing yards for wins and losses
 ggplot(avg_q4_yards_by_outcome, aes(x = as.factor(Win), y = AverageYards, fill = YardType)) +
@@
 
-```{r}
+```{r,fig.cap= "The Fourth-Quarter Rushing and Passing Yards of Every Game in 2023"}
 #| label: fig-5
@@
 
-```{r}
+```{r, fig.cap= "Distribution of Rushing Yards by Wins and Losses"}
 #| label: fig-6
@@
 ```
-```{r}
+```{r, fig.cap= "The Distribution of Passing Yards by Wins and Losses"}
 #| label: fig-7
@@
 
-# 5 Discussion
+# 5 Discussion {#sec-discussion}
 

### 2024-04-02 data_analysis/data.R

@@ -1,0 +1,49 @@
+library(tidyverse)
+library(nflverse)
+
+# Load play-by-play data for a specific season
+pbp_data <- load_pbp(2023) # Adjust the season as needed
+
+# Calculate overall team statistics per game
+team_game_stats <- pbp_data %>%
+  group_by(game_id, posteam) %>%
+  summarize(RushingYards = sum(rushing_yards, na.rm = TRUE),
+            PassingYards = sum(passing_yards, na.rm = TRUE),
+            TotalYards = RushingYards + PassingYards,
+            .groups = 'drop')
+
+# Calculate Q4 team statistics per game
+q4_team_game_stats <- pbp_data %>%
+  filter(qtr == 4) %>%
+  group_by(game_id, posteam) %>%
+  summarize(Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
+            Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
+            Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
+            .groups = 'drop')
+
+# Extract game scores and determine the winner
+game_scores <- pbp_data %>%
+  select(game_id, home_team, away_team, home_score, away_score) %>%
+  distinct(game_id, .keep_all = TRUE) %>%
+  mutate(Winner = case_when(
+    home_score > away_score ~ home_team,
+    away_score > home_score ~ away_team,
+    TRUE ~ "Tie"
+  ))
+
+# Merge overall stats, Q4 stats, and game scores to create a comprehensive dataset
+final_data_with_q4 <- team_game_stats %>%
+  left_join(q4_team_game_stats, by = c("game_id", "posteam")) %>%
+  left_join(game_scores, by = "game_id") %>%
+  mutate(Win = ifelse(posteam == Winner, 1, 0)) %>%
+  select(game_id, posteam, RushingYards, PassingYards, TotalYards, home_score, away_score, Winner, Win, Q4_RushingYards, Q4_PassingYards, Q4_TotalYards)
+
+# Filtering step to remove any rows where 'posteam' might be NA, ensuring cleanliness
+cleaned_data_1 <- final_data_with_q4 %>%
+  filter(!is.na(posteam))
+
+# Display the structure of the cleaned data to confirm column arrangements
+glimpse(cleaned_data)
+
+
+write_csv(cleaned_data, file = "~/nfl_analysis/data/analysis_data/cleaned_data.csv")

### 2024-04-02 data_analysis/data_fun.R

@@ -1,0 +1,63 @@
+library(dplyr)
+
+# Read the dataset
+df <- read.csv('~/nfl_analysis/data/analysis_data/cleaned_data.csv')
+
+# Create a helper function to identify wins with fewer rushing or passing yards
+analyze_wins <- function(df) {
+  # Pair each team with its opponent within the same game
+  df <- df %>%
+    group_by(game_id) %>%
+    mutate(opponent_rushing_yards = lead(RushingYards, order_by = game_id, default = first(RushingYards)),
+           opponent_passing_yards = lead(PassingYards, order_by = game_id, default = first(PassingYards))) %>%
+    ungroup()
+  
+  # Count wins with fewer rushing yards than opponent
+  wins_with_fewer_rushing_yards <- sum(df$Win == 1 & df$RushingYards < df$opponent_rushing_yards, na.rm = TRUE)
+  
+  # Count wins with fewer passing yards than opponent
+  wins_with_fewer_passing_yards <- sum(df$Win == 1 & df$PassingYards < df$opponent_passing_yards, na.rm = TRUE)
+  
+  return(list(wins_with_fewer_rushing_yards = wins_with_fewer_rushing_yards,
+              wins_with_fewer_passing_yards = wins_with_fewer_passing_yards))
+}
+
+# Analyze the data
+results <- analyze_wins(df)
+
+# Print the results
+print(paste("Number of wins with fewer rushing yards than opponent:", results$wins_with_fewer_rushing_yards))
+print(paste("Number of wins with fewer passing yards than opponent:", results$wins_with_fewer_passing_yards))
+
+
+library(dplyr)
+
+# Assuming df is your dataframe loaded from cleaned_data.csv
+# First, create opponent columns for rushing and passing yards to compare within the same game
+df <- df %>%
+  group_by(game_id) %>%
+  mutate(
+    opponent_rushing_yards = if_else(posteam == lead(posteam, order_by = game_id), lag(RushingYards), lead(RushingYards)),
+    opponent_passing_yards = if_else(posteam == lead(posteam, order_by = game_id), lag(PassingYards), lead(PassingYards))
+  ) %>%
+  ungroup()
+
+# Now, filter for conditions where a team won with more rushing AND more passing yards than their opponent
+wins_with_more_rushing_and_passing_yards <- df %>%
+  filter(Win == 1 & RushingYards > opponent_rushing_yards & PassingYards > opponent_passing_yards) %>%
+  nrow()
+
+# Print the result
[…]

### 2024-04-02 data_analysis/gpt_data.R

@@ -1,0 +1,46 @@
+library(tidyverse)
+library(nflreadr)
+library(nflverse)
+
+# Load play-by-play data for a specific season, e.g., 2022
+pbp_data <- load_pbp(2023) # Adjust the season as needed
+view(pbp_data)
+# Calculate team statistics (rushing yards, passing yards, total yards) per game
+team_game_stats <- pbp_data %>%
+  group_by(game_id, posteam) %>%
+  summarize(
+    RushingYards = sum(rushing_yards, na.rm = TRUE),
+    PassingYards = sum(passing_yards, na.rm = TRUE),
+    TotalYards = RushingYards + PassingYards,
+    .groups = 'drop'
+  )
+
+# Assuming home_score and away_score represent final scores, extract these at the game level
+game_scores <- pbp_data %>%
+  select(game_id, home_team, away_team, home_score, away_score) %>%
+  distinct(game_id, .keep_all = TRUE) %>%
+  mutate(Winner = case_when(
+    home_score > away_score ~ home_team,
+    away_score > home_score ~ away_team,
+    TRUE ~ "Tie" # or NA_character_ to indicate a tie differently
+  ))
+
+# Merge the team statistics with the winners and scores to include comprehensive game information
+final_stats_with_winners <- team_game_stats %>%
+  left_join(game_scores, by = "game_id") %>%
+  select(game_id, posteam, RushingYards, PassingYards, TotalYards, home_score, away_score, Winner)
+
+
+# Assuming you have already loaded your play-by-play data into a variable called pbp_data
+# Remove rows where 'posteam' is NA
+cleaned_data <- final_stats_with_winners %>%
+  filter(!is.na(posteam))
+
+final_data_for_modeling <- cleaned_data %>%
+  mutate(Win = ifelse(posteam == Winner, 1, 0))
+# Now, you can proceed with your data analysis or modeling with cleaned_data
+final_data_for_modeling 
+# View the final dataset
+write_csv(x = final_data_for_modeling, file = "~/nfl_analysis/data/analysis_data/gay_cleaned_data.csv")
+
+

### 2024-04-02 data_analysis/gpt_data_4thq.R

@@ -1,0 +1,48 @@
+library(tidyverse)
+library(nflreadr)
+library(nflverse)
+
+# Load play-by-play data for a specific season, e.g., 2023
+pbp_data <- load_pbp(2023)
+
+# Filter for fourth quarter plays only
+q4_data <- pbp_data %>%
+  filter(qtr == 4)
+
+# Calculate team statistics for the fourth quarter (rushing yards, passing yards, total yards) per game
+q4_team_game_stats <- q4_data %>%
+  group_by(game_id, posteam) %>%
+  summarize(
+    Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
+    Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
+    Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
+    .groups = 'drop'
+  )
+
+# Assuming home_score and away_score represent final scores, extract these at the game level from the original dataset
+game_scores <- pbp_data %>%
+  select(game_id, home_team, away_team, home_score, away_score) %>%
+  distinct(game_id, .keep_all = TRUE) %>%
+  mutate(Winner = case_when(
+    home_score > away_score ~ home_team,
+    away_score > home_score ~ away_team,
+    TRUE ~ "Tie" # or NA_character_ to indicate a tie differently
+  ))
+
+# Merge the fourth quarter team statistics with the winners and scores to include comprehensive game information
+q4_final_stats_with_winners <- q4_team_game_stats %>%
+  left_join(game_scores, by = "game_id") %>%
+  select(game_id, posteam, Q4_RushingYards, Q4_PassingYards, Q4_TotalYards, home_score, away_score, Winner)
+
+# Prepare the final data for modeling or analysis, including determining the win/loss status
+q4_final_data_for_modeling <- q4_final_stats_with_winners %>%
+  mutate(Win = ifelse(posteam == Winner, 1, 0))
+
+q4_final_data_for_modeling <- q4_final_data_for_modeling %>%
+  filter(!is.na(posteam))
+
+
+# Output the final dataset for Q4 analysis
+write_csv(q4_final_data_for_modeling, file = "~/nfl_analysis/data/analysis_data/q4_cleaned_data.csv")
+
+# Note: You may need to adjust the file path for 'write_csv' to match your directory structure

### 2024-04-02 data_analysis/gpt_graphs.R

@@ -1,0 +1,136 @@
+library(tidyverse)
+
+ggplot(final_data_for_modeling, aes(x = RushingYards, y = PassingYards, color = as.factor(Win))) +
+  geom_point() +
+  labs(title = "Rushing Yards vs. Passing Yards by Win/Loss",
+       x = "Rushing Yards",
+       y = "Passing Yards",
+       color = "Outcome") +
+  theme_minimal()
+
+ggplot(final_data_for_modeling, aes(x = as.factor(Win), y = TotalYards, fill = as.factor(Win))) +
+  geom_boxplot() +
+  labs(title = "Distribution of Total Yards by Win/Loss",
+       x = "Outcome (0 = Loss, 1 = Win)",
+       y = "Total Yards",
+       fill = "Outcome") +
+  theme_minimal()
+
+ggplot(final_data_for_modeling, aes(x = TotalYards)) +
+  geom_histogram(binwidth = 50, fill = "skyblue", color = "black") +
+  labs(title = "Histogram of Total Yards",
+       x = "Total Yards",
+       y = "Count") +
+  theme_minimal()
+
+final_data_for_modeling %>%
+  group_by(Win) %>%
+  summarize(AvgTotalYards = mean(TotalYards, na.rm = TRUE)) %>%
+  ggplot(aes(x = as.factor(Win), y = AvgTotalYards, fill = as.factor(Win))) +
+  geom_bar(stat = "identity") +
+  labs(title = "Average Total Yards by Win/Loss",
+       x = "Outcome (0 = Loss, 1 = Win)",
+       y = "Average Total Yards",
+       fill = "Outcome") +
+  theme_minimal()
+
+ggplot(final_data_for_modeling, aes(x = RushingYards, fill = as.factor(Win))) +
+  geom_density(alpha = 0.5) +
+  labs(title = "Density of Rushing Yards by Win/Loss",
+       x = "Rushing Yards",
+       y = "Density",
+       fill = "Outcome") +
+  theme_minimal()
+
+ggplot(final_data_for_modeling, aes(x = PassingYards, fill = as.factor(Win))) +
+  geom_density(alpha = 0.5) +
+  labs(title = "Density of Passing Yards by Win/Loss",
+       x = "Rushing Yards",
+       y = "Density",
+       fill = "Outcome") +
+  theme_minimal()
+
+ggplot(final_data_for_modeling, aes(x = RushingYards, y = TotalYards)) +
[…]

### 2024-04-02 data_analysis/data.R
remove unneccesary data exploration files

@@ -1,49 +1,0 @@
-library(tidyverse)
-library(nflverse)
-
-# Load play-by-play data for a specific season
-pbp_data <- load_pbp(2023) # Adjust the season as needed
-
-# Calculate overall team statistics per game
-team_game_stats <- pbp_data %>%
-  group_by(game_id, posteam) %>%
-  summarize(RushingYards = sum(rushing_yards, na.rm = TRUE),
-            PassingYards = sum(passing_yards, na.rm = TRUE),
-            TotalYards = RushingYards + PassingYards,
-            .groups = 'drop')
-
-# Calculate Q4 team statistics per game
-q4_team_game_stats <- pbp_data %>%
-  filter(qtr == 4) %>%
-  group_by(game_id, posteam) %>%
-  summarize(Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
-            Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
-            Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
-            .groups = 'drop')
-
-# Extract game scores and determine the winner
-game_scores <- pbp_data %>%
-  select(game_id, home_team, away_team, home_score, away_score) %>%
-  distinct(game_id, .keep_all = TRUE) %>%
-  mutate(Winner = case_when(
-    home_score > away_score ~ home_team,
-    away_score > home_score ~ away_team,
-    TRUE ~ "Tie"
-  ))
-
-# Merge overall stats, Q4 stats, and game scores to create a comprehensive dataset
-final_data_with_q4 <- team_game_stats %>%
-  left_join(q4_team_game_stats, by = c("game_id", "posteam")) %>%
-  left_join(game_scores, by = "game_id") %>%
-  mutate(Win = ifelse(posteam == Winner, 1, 0)) %>%
-  select(game_id, posteam, RushingYards, PassingYards, TotalYards, home_score, away_score, Winner, Win, Q4_RushingYards, Q4_PassingYards, Q4_TotalYards)
-
-# Filtering step to remove any rows where 'posteam' might be NA, ensuring cleanliness
-cleaned_data_1 <- final_data_with_q4 %>%
-  filter(!is.na(posteam))
-
-# Display the structure of the cleaned data to confirm column arrangements
-glimpse(cleaned_data)
-
-
-write_csv(cleaned_data, file = "~/nfl_analysis/data/analysis_data/cleaned_data.csv")

### 2024-04-02 data_analysis/data_fun.R

@@ -1,63 +1,0 @@
-library(dplyr)
-
-# Read the dataset
-df <- read.csv('~/nfl_analysis/data/analysis_data/cleaned_data.csv')
-
-# Create a helper function to identify wins with fewer rushing or passing yards
-analyze_wins <- function(df) {
-  # Pair each team with its opponent within the same game
-  df <- df %>%
-    group_by(game_id) %>%
-    mutate(opponent_rushing_yards = lead(RushingYards, order_by = game_id, default = first(RushingYards)),
-           opponent_passing_yards = lead(PassingYards, order_by = game_id, default = first(PassingYards))) %>%
-    ungroup()
-  
-  # Count wins with fewer rushing yards than opponent
-  wins_with_fewer_rushing_yards <- sum(df$Win == 1 & df$RushingYards < df$opponent_rushing_yards, na.rm = TRUE)
-  
-  # Count wins with fewer passing yards than opponent
-  wins_with_fewer_passing_yards <- sum(df$Win == 1 & df$PassingYards < df$opponent_passing_yards, na.rm = TRUE)
-  
-  return(list(wins_with_fewer_rushing_yards = wins_with_fewer_rushing_yards,
-              wins_with_fewer_passing_yards = wins_with_fewer_passing_yards))
-}
-
-# Analyze the data
-results <- analyze_wins(df)
-
-# Print the results
-print(paste("Number of wins with fewer rushing yards than opponent:", results$wins_with_fewer_rushing_yards))
-print(paste("Number of wins with fewer passing yards than opponent:", results$wins_with_fewer_passing_yards))
-
-
-library(dplyr)
-
-# Assuming df is your dataframe loaded from cleaned_data.csv
-# First, create opponent columns for rushing and passing yards to compare within the same game
-df <- df %>%
-  group_by(game_id) %>%
-  mutate(
-    opponent_rushing_yards = if_else(posteam == lead(posteam, order_by = game_id), lag(RushingYards), lead(RushingYards)),
-    opponent_passing_yards = if_else(posteam == lead(posteam, order_by = game_id), lag(PassingYards), lead(PassingYards))
-  ) %>%
-  ungroup()
-
-# Now, filter for conditions where a team won with more rushing AND more passing yards than their opponent
-wins_with_more_rushing_and_passing_yards <- df %>%
-  filter(Win == 1 & RushingYards > opponent_rushing_yards & PassingYards > opponent_passing_yards) %>%
-  nrow()
-
-# Print the result
[…]

### 2024-04-02 data_analysis/gpt_data.R

@@ -1,46 +1,0 @@
-library(tidyverse)
-library(nflreadr)
-library(nflverse)
-
-# Load play-by-play data for a specific season, e.g., 2022
-pbp_data <- load_pbp(2023) # Adjust the season as needed
-view(pbp_data)
-# Calculate team statistics (rushing yards, passing yards, total yards) per game
-team_game_stats <- pbp_data %>%
-  group_by(game_id, posteam) %>%
-  summarize(
-    RushingYards = sum(rushing_yards, na.rm = TRUE),
-    PassingYards = sum(passing_yards, na.rm = TRUE),
-    TotalYards = RushingYards + PassingYards,
-    .groups = 'drop'
-  )
-
-# Assuming home_score and away_score represent final scores, extract these at the game level
-game_scores <- pbp_data %>%
-  select(game_id, home_team, away_team, home_score, away_score) %>%
-  distinct(game_id, .keep_all = TRUE) %>%
-  mutate(Winner = case_when(
-    home_score > away_score ~ home_team,
-    away_score > home_score ~ away_team,
-    TRUE ~ "Tie" # or NA_character_ to indicate a tie differently
-  ))
-
-# Merge the team statistics with the winners and scores to include comprehensive game information
-final_stats_with_winners <- team_game_stats %>%
-  left_join(game_scores, by = "game_id") %>%
-  select(game_id, posteam, RushingYards, PassingYards, TotalYards, home_score, away_score, Winner)
-
-
-# Assuming you have already loaded your play-by-play data into a variable called pbp_data
-# Remove rows where 'posteam' is NA
-cleaned_data <- final_stats_with_winners %>%
-  filter(!is.na(posteam))
-
-final_data_for_modeling <- cleaned_data %>%
-  mutate(Win = ifelse(posteam == Winner, 1, 0))
-# Now, you can proceed with your data analysis or modeling with cleaned_data
-final_data_for_modeling 
-# View the final dataset
-write_csv(x = final_data_for_modeling, file = "~/nfl_analysis/data/analysis_data/gay_cleaned_data.csv")
-
-

### 2024-04-02 data_analysis/gpt_data_4thq.R

@@ -1,48 +1,0 @@
-library(tidyverse)
-library(nflreadr)
-library(nflverse)
-
-# Load play-by-play data for a specific season, e.g., 2023
-pbp_data <- load_pbp(2023)
-
-# Filter for fourth quarter plays only
-q4_data <- pbp_data %>%
-  filter(qtr == 4)
-
-# Calculate team statistics for the fourth quarter (rushing yards, passing yards, total yards) per game
-q4_team_game_stats <- q4_data %>%
-  group_by(game_id, posteam) %>%
-  summarize(
-    Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
-    Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
-    Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
-    .groups = 'drop'
-  )
-
-# Assuming home_score and away_score represent final scores, extract these at the game level from the original dataset
-game_scores <- pbp_data %>%
-  select(game_id, home_team, away_team, home_score, away_score) %>%
-  distinct(game_id, .keep_all = TRUE) %>%
-  mutate(Winner = case_when(
-    home_score > away_score ~ home_team,
-    away_score > home_score ~ away_team,
-    TRUE ~ "Tie" # or NA_character_ to indicate a tie differently
-  ))
-
-# Merge the fourth quarter team statistics with the winners and scores to include comprehensive game information
-q4_final_stats_with_winners <- q4_team_game_stats %>%
-  left_join(game_scores, by = "game_id") %>%
-  select(game_id, posteam, Q4_RushingYards, Q4_PassingYards, Q4_TotalYards, home_score, away_score, Winner)
-
-# Prepare the final data for modeling or analysis, including determining the win/loss status
-q4_final_data_for_modeling <- q4_final_stats_with_winners %>%
-  mutate(Win = ifelse(posteam == Winner, 1, 0))
-
-q4_final_data_for_modeling <- q4_final_data_for_modeling %>%
-  filter(!is.na(posteam))
-
-
-# Output the final dataset for Q4 analysis
-write_csv(q4_final_data_for_modeling, file = "~/nfl_analysis/data/analysis_data/q4_cleaned_data.csv")
-
-# Note: You may need to adjust the file path for 'write_csv' to match your directory structure

### 2024-04-02 data_analysis/gpt_graphs.R

@@ -1,136 +1,0 @@
-library(tidyverse)
-
-ggplot(final_data_for_modeling, aes(x = RushingYards, y = PassingYards, color = as.factor(Win))) +
-  geom_point() +
-  labs(title = "Rushing Yards vs. Passing Yards by Win/Loss",
-       x = "Rushing Yards",
-       y = "Passing Yards",
-       color = "Outcome") +
-  theme_minimal()
-
-ggplot(final_data_for_modeling, aes(x = as.factor(Win), y = TotalYards, fill = as.factor(Win))) +
-  geom_boxplot() +
-  labs(title = "Distribution of Total Yards by Win/Loss",
-       x = "Outcome (0 = Loss, 1 = Win)",
-       y = "Total Yards",
-       fill = "Outcome") +
-  theme_minimal()
-
-ggplot(final_data_for_modeling, aes(x = TotalYards)) +
-  geom_histogram(binwidth = 50, fill = "skyblue", color = "black") +
-  labs(title = "Histogram of Total Yards",
-       x = "Total Yards",
-       y = "Count") +
-  theme_minimal()
-
-final_data_for_modeling %>%
-  group_by(Win) %>%
-  summarize(AvgTotalYards = mean(TotalYards, na.rm = TRUE)) %>%
-  ggplot(aes(x = as.factor(Win), y = AvgTotalYards, fill = as.factor(Win))) +
-  geom_bar(stat = "identity") +
-  labs(title = "Average Total Yards by Win/Loss",
-       x = "Outcome (0 = Loss, 1 = Win)",
-       y = "Average Total Yards",
-       fill = "Outcome") +
-  theme_minimal()
-
-ggplot(final_data_for_modeling, aes(x = RushingYards, fill = as.factor(Win))) +
-  geom_density(alpha = 0.5) +
-  labs(title = "Density of Rushing Yards by Win/Loss",
-       x = "Rushing Yards",
-       y = "Density",
-       fill = "Outcome") +
-  theme_minimal()
-
-ggplot(final_data_for_modeling, aes(x = PassingYards, fill = as.factor(Win))) +
-  geom_density(alpha = 0.5) +
-  labs(title = "Density of Passing Yards by Win/Loss",
-       x = "Rushing Yards",
-       y = "Density",
-       fill = "Outcome") +
-  theme_minimal()
-
-ggplot(final_data_for_modeling, aes(x = RushingYards, y = TotalYards)) +
[…]

### 2024-04-02 paper/paper.qmd
Add Sketches

@@
 title: "Field Dynamics" 
-subtitle: "An Analysis of Offensive Strategies in the Modern Day NFL"
+subtitle: "Why the Rush Game is still King in the Modern Day NFL"
 author: "Alexander Guarasci"
@@
        x = "Outcome (0 = Loss, 1 = Win)",
-       y = "Average Yards in Q4",
+       y = "Average Fourth-Quarter Yards",
        fill = "Type of Yards") +
@@
   geom_point() +
-  labs(title = "Rushing Yards vs. Passing Yards by Win/Loss 4th Q",
-       x = "Rushing Yards",
-       y = "Passing Yards",
+  labs(title = "Fourth-Quarter Rushing vs. Passing Yards by Win/Loss",
+       x = "Fourth-Quarter Rushing Yards",
+       y = "Fourth-Quarter Passing Yards",
        color = "Outcome") +
@@
 
-```{r, fig.cap= "Distribution of Rushing Yards by Wins and Losses"}
+```{r, fig.cap= "Distribution of Fourth-Quarter Rushing Yards by Wins and Losses"}
 #| label: fig-6
@@
   geom_density(alpha = 0.5) +
-  labs(title = "Density of Rushing Yards by Win/Loss",
-       x = "Rushing Yards",
+  labs(title = "Density of Fourth-Quarter Rushing Yards by Win/Loss",
+       x = "Fourth-Quarter Rushing Yards",
        y = "Density",
@@
 ```
-```{r, fig.cap= "The Distribution of Passing Yards by Wins and Losses"}
+```{r, fig.cap= "The Distribution of Fourth-Quarter Passing Yards by Wins and Losses"}
 #| label: fig-7
@@
   geom_density(alpha = 0.5) +
-  labs(title = "Density of Passing Yards by Win/Loss",
-       x = "Rushing Yards",
+  labs(title = "Density of Fourth-Quarter Passing Yards by Win/Loss",
+       x = "Fourth-Quarter Rushing Yards",
        y = "Density",

### 2024-04-02 models/model_1.R

@@ -1,0 +1,26 @@
+library(tidyverse)
+library(nflverse)
+library(caret)
+
+df <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")  # Update the path to your CSV file
+
+
+set.seed(10) 
+trainingIndex <- createDataPartition(df$Win, p = .8, list = FALSE)
+train_data <- df[trainingIndex, ]
+test_data <- df[-trainingIndex, ]
+
+
+model_1 <- glm(Win ~ RushingYards + PassingYards, data = train_data, family = binomial())
+
+
+summary(model_1)
+
+
+predictions_prob <- predict(model_1, newdata = test_data, type = "response")
+predictions_class <- ifelse(predictions_prob > 0.5, 1, 0)
+
+
+confusionMatrix(factor(predictions_class), factor(test_data$Win))
+
+save(model_1, confusion_matrix_q4, file = "~/nfl_analysis/models/model_1.RData")

### 2024-04-02 models/model_q4.R

@@ -1,0 +1,26 @@
+library(tidyverse)
+library(caret)
+
+
+df <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")  # Update the path to your CSV file
+
+
+set.seed(10)
+trainingIndex <- createDataPartition(df$Win, p = .8, list = FALSE)
+train_data <- df[trainingIndex, ]
+test_data <- df[-trainingIndex, ]
+
+model_q4 <- glm(Win ~ Q4_RushingYards + Q4_PassingYards, family = binomial(), data = train_data)
+
+predictions_prob <- predict(model_q4, newdata = test_data, type = "response")
+predictions_class <- ifelse(predictions_prob > 0.5, 1, 0)
+
+
+confusion_matrix_q4 <- confusionMatrix(factor(predictions_class), factor(test_data$Win))
+
+
+summary(model_q4)
+confusion_matrix_q4
+
+save(model_q4, confusion_matrix_q4, file = "~/nfl_analysis/models/model_q4.RData")
+

### 2024-04-02 scripts/data_cleaning.R

@@ -1,0 +1,52 @@
+#### Preamble ####
+# Purpose: Cleans and writes the data
+# Author: Alexander Guarasci
+# Date: 6 April 2023 [...UPDATE THIS...]
+# Contact: alexander.guarasci@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: install nflverse and tidyverse
+
+library(tidyverse)
+library(nflverse)
+
+
+pbp_data <- load_pbp(2023)
+
+
+team_game_stats <- pbp_data |>
+  group_by(game_id, posteam) |>
+  summarize(RushingYards = sum(rushing_yards, na.rm = TRUE),
+            PassingYards = sum(passing_yards, na.rm = TRUE),
+            TotalYards = RushingYards + PassingYards,
+            .groups = 'drop')
+
+
+q4_team_game_stats <- pbp_data |>
+  filter(qtr == 4) |>
+  group_by(game_id, posteam) |>
+  summarize(Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
+            Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
+            Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
+            .groups = 'drop')
+
+
+game_scores <- pbp_data |>
+  select(game_id, home_team, away_team, home_score, away_score) |>
+  distinct(game_id, .keep_all = TRUE) |>
+  mutate(Winner = case_when(
+    home_score > away_score ~ home_team,
+    away_score > home_score ~ away_team,
+    TRUE ~ "Tie"
+  ))
+
+final_data_with_q4 <- team_game_stats |>
+  left_join(q4_team_game_stats, by = c("game_id", "posteam")) |>
+  left_join(game_scores, by = "game_id") |>
+  mutate(Win = ifelse(posteam == Winner, 1, 0)) |>
+  select(game_id, posteam, RushingYards, PassingYards, TotalYards, home_score, away_score, Winner, Win, Q4_RushingYards, Q4_PassingYards, Q4_TotalYards)
+
+
+cleaned_data_1 <- final_data_with_q4 |>
+  filter(!is.na(posteam))
+
+write_csv(cleaned_data, file = "~/nfl_analysis/data/analysis_data/cleaned_data.csv")

### 2024-04-02 scripts/simulate_data.R

@@ -1,0 +1,69 @@
+#### Preamble ####
+# Purpose: Simulates Data
+# Author: Alexander Guarasci
+# Date: 6 April 2023 [...UPDATE THIS...]
+# Contact: alexander.guarasci@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: install nflverse and tidyverse
+
+# Simulate some data for NFL games
+set.seed(123) # Ensure reproducibility
+
+# Define the number of games to simulate
+n_games <- 285
+
+# Teams
+teams <- c("ARI", "WAS", "BUF", "NYJ", "CAR", "ATL", "CIN", "CLE", "DAL", "NYG", "DET", "KC", "GB", "CHI", "HOU", "BAL")
+
+# Generate game IDs
+game_ids <- paste(2023, seq_len(n_games), sep="_")
+
+# Simulate data for each game
+simulated_data <- do.call(rbind, lapply(seq_len(n_games), function(i) {
+  home_team <- sample(teams, 1)
+  away_team <- sample(teams[teams != home_team], 1)
+  
+  home_RushingYards <- sample(50:200, 1)
+  away_RushingYards <- sample(50:200, 1)
+  
+  home_PassingYards <- sample(150:400, 1)
+  away_PassingYards <- sample(150:400, 1)
+  
+  home_TotalYards <- home_RushingYards + home_PassingYards
+  away_TotalYards <- away_RushingYards + away_PassingYards
+  
+  home_score <- sample(0:50, 1)
+  away_score <- sample(0:50, 1)
+  
+  Winner <- ifelse(home_score > away_score, home_team, ifelse(home_score < away_score, away_team, "TIE"))
+  Win_home <- ifelse(home_score > away_score, 1, 0)
+  Win_away <- ifelse(away_score > home_score, 1, 0)
+  
+  Q4_home_RushingYards <- sample(20:100, 1)
+  Q4_away_RushingYards <- sample(20:100, 1)
+  
+  Q4_home_PassingYards <- sample(50:200, 1)
+  Q4_away_PassingYards <- sample(50:200, 1)
+  
+  Q4_home_TotalYards <- Q4_home_RushingYards + Q4_home_PassingYards
+  Q4_away_TotalYards <- Q4_away_RushingYards + Q4_away_PassingYards
+  
+  rbind(
+    c(game_id = game_ids[i], posteam = home_team, RushingYards = home_RushingYards, 
+      PassingYards = home_PassingYards, TotalYards = home_TotalYards, home_score = home_score, 
[…]

### 2024-04-02 scripts/test_data.R

@@ -1,0 +1,40 @@
+#### Preamble ####
+# Purpose: Tests to make sure all my data is appropriate
+# Author: Alexander Guarasci
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: alexander.guarasci@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run "data_cleaning.R" 
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+# Load the data
+data <- read_csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+
+# Function to print test results
+print_test_result <- function(test_name, result) {
+  if(result) {
+    message(test_name, ": PASS")
+  } else {
+    message(test_name, ": FAIL")
+  }
+}
+
+# Test 'posteam' and 'Winner' are all 3-character words
+print_test_result("posteam has 3-character words", all(nchar(data$posteam) == 3))
+print_test_result("Winner has 3-character words", all(nchar(data$Winner) == 3))
+
+# Test specific columns are all integers
+integer_columns <- c("RushingYards", "PassingYards", "TotalYards", 
+                     "home_score", "away_score", "Q4_RushingYards", 
+                     "Q4_PassingYards", "Q4_TotalYards")
+
+for (col in integer_columns) {
+  test_result <- all(data[[col]] == floor(data[[col]]))
+  print_test_result(paste(col, "is integers"), test_result)
+}
+
+# Test 'Win' is only 0's and 1's
+print_test_result("Win is only 0's and 1's", all(data$Win %in% c(0, 1)))

### 2024-04-05 paper/paper.qmd
update paper

@@
 ---
-title: "Field Dynamics" 
-subtitle: "Why the Rush Game is still King in the Modern Day NFL"
+title: "Breaking Down the NFL: How Offensive Yards Shape Game Outcomes"
+subtitle: "Examining the Trends in Yardage of Winning NFL Teams Using Logistic Regression"
+
 author: "Alexander Guarasci"
@@
 abstract: |
- This study examines the strategy behind National Football League (NFL) games, focusing on how teams' offensive yardage—gained through both rushing and passing—affects their chances of winning.
+ This study examines the strategy behind National Football League (NFL) games, focusing on how teams' offensive yardage—gained through both rushing and passing—affects their chances of winning using two logistic regression models.
 Analyzing play-by-play data from the 2023 NFL season, we discovered that while both rushing and passing yards contribute to game victories, a stronger rushing attack significantly increases a team's likelihood of winning.
@@
 
-# Introduction {#sec-intro}
+# Introduction {#sec-introduction}
 
@@
 
-This comprehensive dataset encapsulates a wide array of variables, central to which are “RushingYards”, “PassingYards”, “TotalYards”, “home_score”, “away_score”, “Winner”, “Win”, “Q4_RushingYards”, “Q4_PassingYards” and “Q4_TotalYards”.
-RushingYards and  PassingYards represent the yards gained by a team through rushing and passing plays, respectively, within a single game, while Q4_RushingYards and Q4_PassingYards represent the rushing and passing yards in the fourth-quarter.
-“RushingYards” is taken as the sum of the variable “rushing_yards” from the original dataset with the same “game_id” (so that we get the rushing yards from each game) and “posteam” (so each team that plays gets has their rushing yards tracked).
-PassingYards, Q4_RushingYards and Q4_PassingYards are similarly constructed.
-TotalYards is a constructed variable, representing the sum of RushingYards and PassingYards, devised to encapsulate a team's overall offensive performance.
-Q4_TotalYards is constructed as the sum of Q4_RushingYards and Q4_PassingYards.
-The scores of the home and away teams are captured by home_score and away_score, respectively, which is data that is provided and that does not need to be constructed.
-The Winner is determined based on the comparison of these scores, classified as the team that won (for example, “Bal” if the Baltimore Ravens won).
-Finally, the binary Win variable signifies the outcome from the perspective of the team being analyzed (1 for a win, 0 for a loss, and “TIE” for a tie, but because there were no ties last season there are no instances of “TIE”).
+This comprehensive dataset encapsulates a wide array of variables, central to which are “**RushingYards**”, “**PassingYards**”, “**TotalYards**”, “**home_score**”, “**away_score**”, “**Winner**”, “**Win**”, “**Q4_RushingYards**”, “**Q4_PassingYards**” and “**Q4_TotalYards**”.
+**RushingYards** and  **PassingYards** represent the yards gained by a team through rushing and passing plays, respectively, within a single game, while **Q4_RushingYards** and **Q4_PassingYards** represent the rushing and passing yards in the fourth-quarter.
+**RushingYards** is taken as the sum of the variable “**rushing_yards**” from the original dataset with the same “**game_id**” (so that we get the rushing yards from each game) and “**posteam**” (so each team that plays gets has their rushing yards tracked).
+**PassingYards**, **Q4_RushingYards** and **Q4_PassingYards** are similarly constructed.
+**TotalYards** is a constructed variable, representing the sum of **RushingYards** and **PassingYards**, devised to encapsulate a team's overall offensive performance.
+**Q4_TotalYards** is constructed as the sum of **Q4_RushingYards** and **Q4_PassingYards**.
+The scores of the home and away teams are captured by **home_score** and **away_score**, respectively, which is data that is provided and that does not need to be constructed.
+The **Winner** is determined based on the comparison of these scores, classified as the team that won (for example, “Bal” if the Baltimore Ravens won).
+Finally, the binary **Win** variable signifies the outcome from the perspective of the team being analyzed (1 for a win, 0 for a loss, and “TIE” for a tie, but because there were no ties last season there are no instances of “TIE”).
 
@@
 
-The first logistic regression model scrutinizes the entire game, using rushing and passing yards as predictor variables to estimate the odds of winning. This model, expressed as $$
+The first logistic regression model scrutinizes the entire game, using rushing and passing yards as predictor variables to estimate the odds of winning. This first model, @eq-model1, expressed as $$
 \log\left(\frac{p_1}{1 - p_1}\right) = \beta_0 + \beta_1 \cdot \text{RushingYards} + \beta_2 \cdot \text{PassingYards}
-$$
+$$ {#eq-model1}
 , where p1 denotes the probability of winning, facilitates a broad analysis of strategic effectiveness.
@@
 
-The second logistic regression model narrows the focus to the fourth quarter, a critical period where games are often decided.
+The second logistic regression model, @eq-model2, narrows the focus to the fourth quarter, a critical period where games are often decided.
 By analyzing the same predictors but confined to the final quarter, this model, $$
 \log\left(\frac{p_2}{1-p_2}\right) = \beta_3 + \beta_4 \cdot \text{Q4\_RushingYards} + \beta_5 \cdot \text{Q4\_PassingYards}
-$$
+$$ {#eq-model2}
 , offers insights into the tactical shifts that dominate the endgame.

### 2024-04-14 scripts/data_cleaning.R
Change the dates

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans and writes the data
 # Author: Alexander Guarasci
-# Date: 6 April 2023 [...UPDATE THIS...]
+# Date: 18 April, 2024
 # Contact: alexander.guarasci@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: install nflverse and tidyverse
@@ -15,19 +15,23 @@ pbp_data <- load_pbp(2023)
 
 team_game_stats <- pbp_data |>
   group_by(game_id, posteam) |>
-  summarize(RushingYards = sum(rushing_yards, na.rm = TRUE),
-            PassingYards = sum(passing_yards, na.rm = TRUE),
-            TotalYards = RushingYards + PassingYards,
-            .groups = 'drop')
+  summarize(
+    RushingYards = sum(rushing_yards, na.rm = TRUE),
+    PassingYards = sum(passing_yards, na.rm = TRUE),
+    TotalYards = RushingYards + PassingYards,
+    .groups = "drop"
+  )
 
 
 q4_team_game_stats <- pbp_data |>
   filter(qtr == 4) |>
   group_by(game_id, posteam) |>
-  summarize(Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
-            Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
-            Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
-            .groups = 'drop')
+  summarize(
+    Q4_RushingYards = sum(rushing_yards, na.rm = TRUE),
+    Q4_PassingYards = sum(passing_yards, na.rm = TRUE),
+    Q4_TotalYards = Q4_RushingYards + Q4_PassingYards,
+    .groups = "drop"
+  )
 
 
 game_scores <- pbp_data |>
@@ -46,7 +50,7 @@ final_data_with_q4 <- team_game_stats |>
   select(game_id, posteam, RushingYards, PassingYards, TotalYards, home_score, away_score, Winner, Win, Q4_RushingYards, Q4_PassingYards, Q4_TotalYards)
 
 
-cleaned_data_1 <- final_data_with_q4 |>
+cleaned_data <- final_data_with_q4 |>
   filter(!is.na(posteam))
 
 write_csv(cleaned_data, file = "~/nfl_analysis/data/analysis_data/cleaned_data.csv")

### 2024-04-14 scripts/simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates Data
 # Author: Alexander Guarasci
-# Date: 6 April 2023 [...UPDATE THIS...]
+# Date: 18 April, 2024
 # Contact: alexander.guarasci@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: install nflverse and tidyverse
@@ -16,48 +16,51 @@ n_games <- 285
 teams <- c("ARI", "WAS", "BUF", "NYJ", "CAR", "ATL", "CIN", "CLE", "DAL", "NYG", "DET", "KC", "GB", "CHI", "HOU", "BAL")
 
 # Generate game IDs
-game_ids <- paste(2023, seq_len(n_games), sep="_")
+game_ids <- paste(2023, seq_len(n_games), sep = "_")
 
 # Simulate data for each game
 simulated_data <- do.call(rbind, lapply(seq_len(n_games), function(i) {
   home_team <- sample(teams, 1)
   away_team <- sample(teams[teams != home_team], 1)
-  
+
   home_RushingYards <- sample(50:200, 1)
   away_RushingYards <- sample(50:200, 1)
-  
+
   home_PassingYards <- sample(150:400, 1)
   away_PassingYards <- sample(150:400, 1)
-  
+
   home_TotalYards <- home_RushingYards + home_PassingYards
   away_TotalYards <- away_RushingYards + away_PassingYards
-  
+
   home_score <- sample(0:50, 1)
   away_score <- sample(0:50, 1)
-  
+
   Winner <- ifelse(home_score > away_score, home_team, ifelse(home_score < away_score, away_team, "TIE"))
   Win_home <- ifelse(home_score > away_score, 1, 0)
   Win_away <- ifelse(away_score > home_score, 1, 0)
-  
+
   Q4_home_RushingYards <- sample(20:100, 1)
   Q4_away_RushingYards <- sample(20:100, 1)
-  
+
   Q4_home_PassingYards <- sample(50:200, 1)
   Q4_away_PassingYards <- sample(50:200, 1)
-  
+
   Q4_home_TotalYards <- Q4_home_RushingYards + Q4_home_PassingYards
   Q4_away_TotalYards <- Q4_away_RushingYards + Q4_away_PassingYards
-  
+
   rbind(
-    c(game_id = game_ids[i], posteam = home_team, RushingYards = home_RushingYards, 
-      PassingYards = home_PassingYards, TotalYards = home_TotalYards, home_score = home_score, 
[…]

### 2024-04-14 scripts/test_data.R

@@ -1,10 +1,10 @@
 #### Preamble ####
 # Purpose: Tests to make sure all my data is appropriate
 # Author: Alexander Guarasci
-# Date: 11 February 2023 [...UPDATE THIS...]
+# Date: 18 April, 2024
 # Contact: alexander.guarasci@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Run "data_cleaning.R" 
+# Pre-requisites: Run "data_cleaning.R"
 
 
 #### Workspace setup ####
@@ -15,7 +15,7 @@ data <- read_csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
 
 # Function to print test results
 print_test_result <- function(test_name, result) {
-  if(result) {
+  if (result) {
     message(test_name, ": PASS")
   } else {
     message(test_name, ": FAIL")
@@ -27,9 +27,11 @@ print_test_result("posteam has 3-character words", all(nchar(data$posteam) == 3)
 print_test_result("Winner has 3-character words", all(nchar(data$Winner) == 3))
 
 # Test specific columns are all integers
-integer_columns <- c("RushingYards", "PassingYards", "TotalYards", 
-                     "home_score", "away_score", "Q4_RushingYards", 
-                     "Q4_PassingYards", "Q4_TotalYards")
+integer_columns <- c(
+  "RushingYards", "PassingYards", "TotalYards",
+  "home_score", "away_score", "Q4_RushingYards",
+  "Q4_PassingYards", "Q4_TotalYards"
+)
 
 for (col in integer_columns) {
   test_result <- all(data[[col]] == floor(data[[col]]))

### 2024-04-14 paper/datasheet.qmd
Add data sheet

---
title: "Datasheet for 'NFLVerse Play-by-Play Data from 2023'"

author: 
  - Alexander Guarasci
thanks: "Code and data are available at: https://github.com/AlexanderG123/nfl_analysis."
date: "April 18, 2024"
format: pdf
number-sections: true
bibliography: references.bib
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*

The dataset was created to analyze play-by-play data of NFL teams.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*

The nflverse authors created the dataset on there own behalf

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*

Could not find information on this, the data was taken from the NFL.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*

The instances that comprise the dataset primarily represent an individual NFL snap, however, there are data on overall game stats (who won, etc). For each instance, there are multiple types, for example, there are stats on points scored, yards gained, where the ball is on the field, the expected points added of a play, and lots more.

2. *How many instances are there in total (of each type, if appropriate)?*

Just under 19 million

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*

It contains all possible instances.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*

Each instance consists of numeric and text data, all of which has been formatted apropriately

5. *Is there a label or target associated with each instance? If so, please provide a description.*
No

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
No

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*

No, the relationships are not made explicit

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*

There is no recommended data splits

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*

There are some redundancies, for example they have data on how many rushing yards and passing yards were gained on a specific play, however, if the ball was run it could not have been passed as well.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*

The dataset is self-contained.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
No

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*

No

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*

No

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*

Yes, this dataset contains the names of individuals who were involved in certain plays.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*

No

16. *Any other comments?*

No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*

The data is directly observable and reported by NFL referees and broadcasters. They also use technology, according to the NFL, they use “20–30 ultra-wide band receiversm, 2–3 radio-frequency identification (RFID) tags installed into the players’ shoulder pads, RFID tags on officials, pylons, sticks, chains, and in the ball. Altogether, an estimated 250 devices are in a venue for any given game.”

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*

As stated above, the NFL uses an estimated 250 devices each game that they play, to validate these procedures,  “a team of three operators is required at every game to confirm that all tracking systems are functioning properly.”

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*

NA

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*

Thousands of people were involved in the data collection process, all of whom were compensated differently, for example, the referees make approximately $12000 for each game they participate in.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*

This data was collected over the 2023 NFL season, the time frame does match the creation timeframe of the data associated with the instances.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*

No

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*

No

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
NA, working as a professional athlete requires that stats are kept while you are employed

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*

NA

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*

NA

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*

NA

12. *Any other comments?*

No

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*

This is a well-known dataset that has likely been used for a wealth of tasks

2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*

No

3. *What (other) tasks could the dataset be used for?*

An analysis of how EPA impacts a team's likelihood of winning could be interesting. There are literally an infinite number of things that can be done with this data. 
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*

No

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*

Because the data is play-by-play, in other words, it documents these data points for each snap, it would be unwise to use it for macro NFL trends, unless they require this granularity.

6. *Any other comments?*

No

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*

No

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*

NA

3. *When will the dataset be distributed?*

NA

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*

NA
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
NA

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*

No

7. *Any other comments?*

No

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*

The kind folks at NFLVerse

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*

https://twitter.com/mrcaseb

3. *Is there an erratum? If so, please provide a link or other access point.*

NA

4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*

NA

5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*

NA
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*

Yes, they are all hosted by NFLVerse

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumewrs? If so, please provide a description.*

NA

8. *Any other comments?*

No

\newpage

# References

### 2024-04-14 paper/paper.qmd
Update Paper

@@
 knitr::opts_chunk$set(echo = FALSE, results = "hide", message = FALSE, warning = FALSE)
-
 ```
@@
 
-This analysis was conducted through the R programming language [@citeR] and supplemented with the analytical tools Tidyverse [@citetidyverse] and caret [@citecaret].
+This analysis was conducted through the R programming language [@citeR] and supplemented with the analytical tools Tidyverse [@citetidyverse], caret [@citecaret] with the code styled using Styler [@citestyler].
 The dataset was assembled from play-by-play records of NFL games from NFLVerse [@citenflverse].
@@
   geom_point() +
-  labs(title = "Rushing Yards vs. Passing Yards by Win/Loss",
+  labs(
+    title = "Rushing Yards vs. Passing Yards by Win/Loss",
     x = "Rushing Yards",
     y = "Passing Yards",
-       color = "Outcome") +
+    color = "Outcome"
+  ) +
   theme_minimal()
@@
 ggplot(data, aes(x = as.factor(Win), y = TotalYards, fill = as.factor(Win))) +
-  geom_violin()+
-  labs(title = "Distribution of Total Yards by Win/Loss",
+  geom_violin() +
+  labs(
+    title = "Distribution of Total Yards by Win/Loss",
     x = "Outcome (0 = Loss, 1 = Win)",
     y = "Total Yards",
-       fill = "Outcome") +
+    fill = "Outcome"
+  ) +
   theme_minimal()
@@
   group_by(Win) |>
-  summarize(AvgRushingYards = mean(RushingYards, na.rm = TRUE),
-            AvgPassingYards = mean(PassingYards, na.rm = TRUE)) |>
+  summarize(
+    AvgRushingYards = mean(RushingYards, na.rm = TRUE),
+    AvgPassingYards = mean(PassingYards, na.rm = TRUE)
+  ) |>
   pivot_longer(cols = -Win, names_to = "YardType", values_to = "AverageYards")
@@
   scale_fill_manual(values = c("AvgRushingYards" = "skyblue", "AvgPassingYards" = "orange")) +
-  labs(title = "Average Rushing and Passing Yards for Wins and Losses",
+  labs(
+    title = "Average Rushing and Passing Yards for Wins and Losses",
     x = "Outcome (0 = Loss, 1 = Win)",
     y = "Average Yards",
-       fill = "Type of Yards") +
+    fill = "Type of Yards"
+  ) +
   theme_minimal()
@@
   group_by(Win) %>%
-  summarize(AvgRushingYards = mean(Q4_RushingYards, na.rm = TRUE),
-            AvgPassingYards = mean(Q4_PassingYards, na.rm = TRUE)) %>%
+  summarize(
+    AvgRushingYards = mean(Q4_RushingYards, na.rm = TRUE),
+    AvgPassingYards = mean(Q4_PassingYards, na.rm = TRUE)
+  ) %>%
   pivot_longer(cols = -Win, names_to = "YardType", values_to = "AverageYards")
@@
   scale_fill_manual(values = c("AvgRushingYards" = "skyblue", "AvgPassingYards" = "orange")) +
-  labs(title = "Average Q4 Rushing and Passing Yards for Wins and Losses",
+  labs(
+    title = "Average Q4 Rushing and Passing Yards for Wins and Losses",
     x = "Outcome (0 = Loss, 1 = Win)",
     y = "Average Fourth-Quarter Yards",
-       fill = "Type of Yards") +
+    fill = "Type of Yards"
+  ) +
   theme_minimal()
@@
 q4_final_data_for_modeling <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
-#4th q data Q4_RushingYards,Q4_PassingYards
+# 4th q data Q4_RushingYards,Q4_PassingYards
 ggplot(q4_final_data_for_modeling, aes(x = Q4_RushingYards, y = Q4_PassingYards, color = as.factor(Win))) +
   geom_point() +
-  labs(title = "Fourth-Quarter Rushing vs. Passing Yards by Win/Loss",
+  labs(
+    title = "Fourth-Quarter Rushing vs. Passing Yards by Win/Loss",
     x = "Fourth-Quarter Rushing Yards",
     y = "Fourth-Quarter Passing Yards",
-       color = "Outcome") +
+    color = "Outcome"
+  ) +
   theme_minimal()
-
 ```
@@
   geom_density(alpha = 0.5) +
-  labs(title = "Density of Fourth-Quarter Rushing Yards by Win/Loss",
+  labs(
+    title = "Density of Fourth-Quarter Rushing Yards by Win/Loss",
     x = "Fourth-Quarter Rushing Yards",
     y = "Density",
-       fill = "Outcome") +
+    fill = "Outcome"
+  ) +
   theme_minimal()
@@
   geom_density(alpha = 0.5) +
-  labs(title = "Density of Fourth-Quarter Passing Yards by Win/Loss",
+  labs(
+    title = "Density of Fourth-Quarter Passing Yards by Win/Loss",
     x = "Fourth-Quarter Rushing Yards",
     y = "Density",
-       fill = "Outcome") +
+    fill = "Outcome"
+  ) +
   theme_minimal()

### 2024-04-17 paper/paper.qmd
Add files via upload
some quick edits to paper.qmd

@@
 Ultimately trying to answer the questions, how do offensive yards impact the outcome of an NFL game, and what is more valuable passing yards or rushing yards[^note]?
+Prior to examining the data, it was our estimation that in the modern, pass friendly NFL, a focus on throwing the football over the course of the game with a switch to the run game in the fourth-quarter is ultimately the best way for teams to win.
 
-[^note]: Rushing yards and running yards are used interchangeably throughout this paper
+[^note]: Rushing yards and running yards are used interchangeably throughout this paper.
 
@@
 
+This study aims to make a substantial contribution to the conversation about sports analytics by offering a strong analytical framework for examining the connection between NFL game-winning and yards.
+It reveals insights into offensive strategy through in-depth statistical analysis and model-building, providing information for analysts, players and fans.
+The results of this study show how important data analysis is to sports strategy and how quantitative methods can potentially help improve competitive advantages in professional football.
+This paper is broken up into four sections: the data section which discusses the data used in this paper; the model section explains the models that were used in this paper as well as why they were appropriate; the results section which explains what information was gained through this analysis; and finally the discussion section discusses the findings as a whole.
+
+# 2 Data {#sec-data}
+
+The data used in this paper was taken from nflverse [@citenflverse], which consists of “a set of packages dedicated to data of the National Football League.”
+
 This analysis was conducted through the R programming language [@citeR] and supplemented with the analytical tools Tidyverse [@citetidyverse], caret [@citecaret] with the code styled using Styler [@citestyler].
@@
 
-This study aims to make a substantial contribution to the conversation about sports analytics by offering a strong analytical framework for examining the connection between NFL game-winning and yards.
-It reveals insights into offensive strategy through in-depth statistical analysis and model-building, providing information for analysts, players and fans.
-The results of this study show how important data analysis is to sports strategy and how quantitative methods can potentially help improve competitive advantages in professional football.
-
-# 2 Data {#sec-data}
-
-The data used in this paper was taken from nflverse [@citenflverse], which consists of “a set of packages dedicated to data of the National Football League.”
-
 # 2.1 The Dataset
@@
 This includes an unbelievably dense array of information from every snap over the entirety of the season consisting of 372 variables.
-These include: seconds left in the half, touchdown probability, expected points added, the probability of taking a safety, the position the snap was taken in, a description of the play, and many, many more.
+These include seconds left in the half, touchdown probability, expected points added, the probability of taking a safety, the position the snap was taken in, a description of the play, and many, many more.
 In order to filter this information the play-by-play data was cleaned and new variables were created.
+
+The papers data preparation and cleaning process was streamlined, because of how robust the data from NFLVerse [@citenflverse] is.
+All that needed to be done was the construction of variables (discussed in section 2.2), but otherwise, there was no missing data or impactful decisions that needed to be made during the preparation phase.
+One thing to note is that the uncleaned dataset is not saved in this repo, this is because the size of the data is too large to store on github so this step was skipped.
 
@@
 , offers insights into the tactical shifts that dominate the endgame.
-Unlike the overall game model, the fourth-quarter model might reveal a nuanced strategic landscape, such as a negative coefficient for passing yards, indicating the potential pitfalls of a pass-heavy approach under pressure.
+Unlike the overall game model, the fourth-quarter model reveals a nuanced strategic landscape, such as a negative coefficient for passing yards, indicating the potential pitfalls of a pass-heavy approach under pressure.
 This focused analysis is instrumental in understanding the dynamics of clutch gameplay, shedding light on the adjustments teams make in response to the scoreboard and the clock.
@@
 The interpretability of logistic regression emphasizes its suitability even more; the exponential of the coefficients accurately measures the change in odds that come with every extra yard gained, whether by passing or rushing.
-This gives you a concise, useful grasp of how various aspects of the game affect your chances of winning.
+This gives a concise, useful grasp of how various aspects of the game affect a team’s chances of winning.
 
@@
 library(tidyverse)
-data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+data <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 ggplot(data, aes(x = RushingYards, y = PassingYards, color = as.factor(Win))) +
@@
 #| label: fig-2
-data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+data <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 ggplot(data, aes(x = as.factor(Win), y = TotalYards, fill = as.factor(Win))) +
@@
 #| label: fig-3
-data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+data <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 avg_yards_by_outcome <- data |>
@@
 #| label: fig-4
-avg_q4_yards_by_outcome <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv") %>%
+avg_q4_yards_by_outcome <- read.csv(here::here("data/analysis_data/cleaned_data.csv")) %>%
   group_by(Win) %>%
@@
 #| label: fig-5
-q4_final_data_for_modeling <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+q4_final_data_for_modeling <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 # 4th q data Q4_RushingYards,Q4_PassingYards
@@
 #| label: fig-6
-data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+data <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 ggplot(data, aes(x = Q4_RushingYards, fill = as.factor(Win))) +
@@
 #| label: fig-7
-data <- read.csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+data <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 ggplot(data, aes(x = Q4_PassingYards, fill = as.factor(Win))) +

### 2024-04-17 scripts/data_cleaning.R
Edit scripts to change how the data is accessed

@@ -53,4 +53,4 @@ final_data_with_q4 <- team_game_stats |>
 cleaned_data <- final_data_with_q4 |>
   filter(!is.na(posteam))
 
-write_csv(cleaned_data, file = "~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+write_csv(cleaned_data, file = here::here("data/analysis_data/cleaned_data.csv"))

### 2024-04-17 scripts/simulate_data.R

@@ -69,4 +69,11 @@ simulated_data_df <- as.data.frame(simulated_data, stringsAsFactors = FALSE)
 numerical_columns <- c("RushingYards", "PassingYards", "TotalYards", "home_score", "away_score", "Win", "Q4_RushingYards", "Q4_PassingYards", "Q4_TotalYards")
 simulated_data_df[numerical_columns] <- lapply(simulated_data_df[numerical_columns], as.numeric)
 
-write_csv(simulated_data_df, file = "~/nfl_analysis/data/analysis_data/sim_data.csv")
+write_csv(simulated_data_df, file = here::here("data/analysis_data/sim_data.csv"))
+          
+          
+          
+          
+          
+          
+          

### 2024-04-17 scripts/test_data.R

@@ -11,7 +11,7 @@
 library(tidyverse)
 
 # Load the data
-data <- read_csv("~/nfl_analysis/data/analysis_data/cleaned_data.csv")
+data <- read.csv(here::here("data/analysis_data/cleaned_data.csv"))
 
 # Function to print test results
 print_test_result <- function(test_name, result) {

### 2024-04-18 scripts/data_cleaning.R
Update data_cleaning.R with comments

@@ -9,10 +9,10 @@
 library(tidyverse)
 library(nflverse)
 
-
+# load dta
 pbp_data <- load_pbp(2023)
 
-
+# construct variiables for rushing yards, passing yards and total yards
 team_game_stats <- pbp_data |>
   group_by(game_id, posteam) |>
   summarize(
@@ -22,7 +22,7 @@ team_game_stats <- pbp_data |>
     .groups = "drop"
   )
 
-
+# construct variiables for q4 rushing yards, passing yards and total yards
 q4_team_game_stats <- pbp_data |>
   filter(qtr == 4) |>
   group_by(game_id, posteam) |>
@@ -33,7 +33,7 @@ q4_team_game_stats <- pbp_data |>
     .groups = "drop"
   )
 
-
+# construct variiables for winner
 game_scores <- pbp_data |>
   select(game_id, home_team, away_team, home_score, away_score) |>
   distinct(game_id, .keep_all = TRUE) |>
@@ -43,6 +43,7 @@ game_scores <- pbp_data |>
     TRUE ~ "Tie"
   ))
 
+# put it all together
 final_data_with_q4 <- team_game_stats |>
   left_join(q4_team_game_stats, by = c("game_id", "posteam")) |>
   left_join(game_scores, by = "game_id") |>
